### 2020-09-30 serial.tex
miscellaneous edits

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|l|l|r|}
\hline
Parameter & Values & Source\\
\hline
Mean intrinsic incubation period & 5.5 days & \cite{lauer2020incubation} \\
SD intrinsic incubation period & 2.4 days & \cite{lauer2020incubation} \\
Mean intrinsic generation interval & 5 days & \cite{ferretti2020quantifying} \\
SD intrinsic generation interval & 2 days & \cite{ferretti2020quantifying} \\
\hline
\end{tabular}
\end{center}
\caption{
  \textbf{Parameter values used for simulations.}
The intrinsic generation-interval distribution is parameterized using a log-normal distribution with log mean $\mu_G=1.54$ and log standard deviation $\sigma_G=0.37$.
The intrinsic incubation period distribution is parameterized using a log-normal distribution with log mean $\mu_I=1.62$ and log standard deviation $\sigma_I=0.42$.
The joint probability distribution is modeled using a bivariate log-normal distribution with correlations (on the log scale) $\rho=0, 0.25, 0.5, 0.75$.
The intrinsic generation-interval and incubation period distributions are chosen to match characteristic of COVID-19 to illustrate realistic magnitudes of time-varying/perspective effects in the current pandemic.
}
\end{table}

@@

We use parameter estimates for COVID-19 to characterize the degree to which the realized serial-interval distribution can change over the course of an epidemic and evaluate how different definitions of the serial-interval distribution can affect the Euler-Lotka estimates of \Ro.
We further address how the observed serial intervals, measured through contact tracing, are affected by the right censoring during an ongoing epidemic and provide a heuristic method for addressing biases that can arise from using serial-interval data to estimate \Ro.
Finally, we analyze serial-interval data from the COVID-19 epidemic in China under our framework.

@@

\fref{rR} shows Euler-Lotka estimates of \Ro\ based on different definitions of the serial interval. 
When the initial forward serial-interval distributions $f_0(\tau)$ is used, estimates (from \eref{Rforward}) exactly match the (correct) generation-interval-based estimates (\eref{Rgen}) for all values of the correlation $\rho$ between the intrinsic incubation period and the intrinsic generation interval (\fref{rR}A).
When the intrinsic distributions are used, however, estimates based on the serial interval (\eref{Rintrinsic}) underestimate \Ro: as $r$ increases, \Rintrinsic saturates and eventually \emph{decreases} due to the increasing inferred importance of negative serial intervals (\fref{rR}B).
While the initial forward serial intervals during the exponential growth phase can also be negative, their effects are appropriately balanced because faster epidemic growth leads to longer serial intervals (and a corresponding lower proportion of negative intervals).

\begin{figure}[!th]
\includegraphics[width=\textwidth]{rR.pdf}
\caption{
\textbf{Estimates of the reproduction number from the exponential growth rate based on serial- and generation-interval distributions.}
(A). The initial forward serial-interval distributions give a correct
link between the exponential growth rate $r$ and the reproduction
number \Ro,
for any correlation $\rho$ between intrinsic incubation period and
intrinsic generation interval of the underlying bivariate log-normal distribution.
(B) The intrinsic serial-interval distributions give an incorrect link between $r$ and \Ro.
(C) The mean initial forward serial interval during the exponential growth phase increases with $r$.
(D) The squared coefficient of variation of the initial forward serial intervals during the exponential growth phase decreases with $r$.
}
\label{fig:rR}
\end{figure}

Comparing the shapes of the initial forward serial-interval distribution (\eref{initialSI}) and the intrinsic generation-interval distribution allows us to better understand how different forward distributions lead to identical estimates of \Ro.
In general, distributions with higher means and less variability lead to higher \Ro for a given $r$ \citep{wallinga2007generation, weitz2015modeling, park2019practical}.
When incidence is growing exponentially, forward serial intervals have higher means (\fref{rR}C) and squared coefficients of variation (\fref{rR}D) than the intrinsic generation-interval distribution.
The effects of higher means (which increase \Ro) exactly cancel those of higher variability (which decrease \Ro).
On the other hand, \emph{intrinsic} serial intervals (\eref{intrinsicSI}) have the same mean (equal to the mean initial forward serial at $r=0$ in \fref{rR}C) as the intrinsic generation intervals but are more variable (also see squared coefficient of variation of the initial forward serials at $r=0$ in \fref{rR}D); 
therefore, we underestimate \Ro when we use the intrinsic serial-interval distribution.

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=0.95\textwidth]{forward.pdf}
\caption{
\textbf{Epidemiological dynamics and changes in mean forward and backward delay distributions.}
(A) Daily incidence over time.
(B--D) Changes in the mean forward incubation period, generation interval, and serial interval.
(E--G) Changes in the mean backward incubation period, generation interval, and serial interval.
Black lines represent the results of a deterministic simulation.
Orange lines (A) represent the results of 10 stochastic simulations.
Orange points (B-G) represent the average of 10 stochastic simulations.
Dashed lines represent the mean initial forward delay.
In order to remove possible transient dynamics (e.g., left-censoring of time delays and initial stochasticity due to low number of infections), we set $t=0$ to the first time point when daily incidence is greater than 100.
Intrinsic incubation periods and intrinsic generation intervals are assumed to be independent of each other for simplicity; simulations with correlated intrinsic incubation periods and intrinsic generation intervals are provided in Supplementary Materials.
See Table 1 for parameter values.
}
\label{fig:epi}
\end{center}
\end{figure}

@@

We see similar qualitative patterns in all three backward delays (\fref{epi}E--G; \eref{backward}), because they are predominantly driven by the rate of change in incidence, which in turn affects relative cohort sizes.
When incidence is increasing, individuals are more likely to have been infected recently and therefore, we are more likely to observe shorter intervals (\eref{backexp}).
Similarly, when incidence decreases, we are more likely to observe longer intervals.
Neglecting these changes will bias the inference of intrinsic distributions from observed distributions.

@@

While the qualitative changes in the mean forward and backward serial interval are consistent with our earlier simulations (\fref{epi}), the initial mean forward serial interval (\fref{du}B) appears to be larger than what we calculated based on previously estimated incubation period and generation-interval distributions (\fref{rR}C).
These may imply that the incubation period and generation interval (Table 1) were underestimated, as neither study explicitly accounted for the fact that the observed intervals were drawn from the backward distributions and were likely to have been censored.

\fref{du}D shows the cohort-averaged estimates of \Ro, which remain roughly constant until day January 17th and suddenly decreases;
this sudden decrease is due changes in the forward serial intervals consistent with the dynamics seen in our simulations (\fref{obsrR}).
The cohort-averaged estimates of \Ro based on the early forward serial intervals are also consistent with previous estimates of \Ro of the COVID-19 epidemic in China \citep{majumder2020early, doi:10.1098/rsif.2020.0144}:
$\Ro = 2.6$ (95\% CI: 2.2--3.1) and $\Ro = 3.4$ (95\% CI: 2.7 -- 4.3) based on a doubling period of 8 or 6 days, respectively, using serial-interval data from infectors who developed symptoms by January 17th.
These early cohort-averaged estimates of \Ro are unlikely to be affected by the right-censoring as we expect the degree of right-censoring to be low (\fref{du}A).
Therefore, the original \Ro estimate of 1.32 (95\% CI 1.16-–1.48), which neglects the changes in the forward serial-interval distribution, underestimates \Ro by 50\%--60\%.
This example demonstrates the danger of using the observed serial intervals to calculate the reproduction number without organizing serial intervals into cohorts.

@@

Recently, \cite{ali2020serial} also showed that serial intervals of COVID-19 decreased in China. Analogous to our approach, they also grouped serial intervals by the symptom onset date of infectors across 14-day periods and found that the mean forward serial interval decreased from 7.8 days to 2.6 days.
While they attributed the decrease in serial intervals to reduction of the isolation delay, their regression analysis showed that isolation delays explain only 51.5\% of the variation in serial intervals (and explaining up to 72\% of the variance including other intervention measures).
Our framework provides an explanation for the remaining variation: changes in the backward incubation period during the decreasing phase of an epidemic also shorten serial intervals due to increased amount of presymptomatic transmission (even in the absence of nonpharmaceutical interventions).
Since isolation delays and other intervention measures affect the amount of onward transmission, and therefore the distribution of realize (forward) generation intervals, it cannot explain all variation in forward serial intervals, which depend on (i) backward incubation period of the infector, (ii) the forward generation interval, and (iii) the forward incubation period of the infectee (\fref{diagram}B).

Our results support the use of serial-interval distributions for calculating the initial \RR during the exponential growth phase,  
but they also reveal gaps in current practices in incorporating serial-interval distributions into outbreak analyses.
For example, \cite{thompson2019improved} recently emphasized the importance of using up-to-date serial-interval data for accurate estimation of time-varying reproduction numbers.
However, our results show that if changes the forward serial interval through time are not accounted for, using up-to-date serial-interval data can, in fact, exacerbate the underestimation of initial \RR.
Future studies should explore how neglecting changes in the forward serial-interval distribution can affect the estimates of \RR beyond the exponential growth phase and potentially re-assess existing estimates of \RR.
We also suggest that modelers should aim to characterize spatiotemporal variation in forward serial-interval distributions and understand their changes over time.
These modeling approaches should be coupled with epidemiological investigation through contact tracing. 
Going forward, an additional advantage of early, intensive contact tracing of emerging diseases is that this is the best way to characterize the initial forward serial-interval distribution.

Our study also underlines the fact that the serial-interval distribution depends not only on the generation-interval and incubation-period distributions, but also on their intrinsic correlations.
Here, we use a bivariate lognormal distribution to capture their correlations phenomenologically but their true correlations will depend on viral load dynamics;
nonetheless, we expect our conclusions about decreasing serial intervals of COVID-19 to be robust as individuals with longer incubation periods have longer time window to transmit before symptom onset.
Otherwise, the impact of increasing backward incubation periods on the forward serial intervals are likely to be disease specific---for example, we show in Supplementary Materials that the initial forward serial-interval distribution can be equivalent to the intrinsic generation-interval distribution, regardless of the growth rate $r$, due to independence between the incubation period and time from symptom onset to transmission (assuming no transmission occurs during the incubation period).
This implies that not considering this correlation can potentially bias the estimates of the intrinsic generation-interval distribution from the serial-interval distribution.
% Previous studies that tried to estimate the generation-interval distributions from the observed serial intervals often ignored the dynamical differences between the realized incubation periods of infectors (backward-looking) and those of infectees (forward looking) (e.g., \cite{klinkenberg2011correlation, ganyani2020estimating}).
Better statistical tools for teasing apart the intrinsic generation-interval distributions from the observed serial intervals are needed.

Here, we lay out a few practical principles for analyzing and interpreting serial-interval data.
First, serial intervals should be cohorted based on the symptom onset date whenever possible.
Previous studies have often regarded serial intervals as an intrinsic quantity, having the same mean as the intrinsic generation interval \citep{svensson2007note,klinkenberg2011correlation,champredon2018equivalence, britton2019estimation}, but the distribution (and the mean) of observed serial intervals is expected to differ from this expectation, and to change through time due to epidemic dynamics.
Second, aggregating serial intervals across different cohorts and epidemic periods should be avoided as they can be subject to different censoring and epidemiological biases:
Even when \emph{all} realized serial intervals can be observed throughout an unmitigated epidemic, we do not obtain the intrinsic serial interval distribution due to susceptible depletion (\fref{obsrR}).
Third, applying serial interval information from one epidemic of a given disease to another should be done with care, because serial intervals are epidemic-specific, rather than disease-specific.
Finally, serial-interval data should be accompanied by a trajectory of the epidemic curve, whenever possible, to provide epidemiological context.
In practice, these recommendations will sometimes be hard to follow, due to limited data about serial intervals, but these issues should be kept in mind when interpreting serial-interval data and understand transmission dynamics at the individual-level.

More broadly, our study underlines the importance of carefully defining measured epidemiological time distributions. 
Previous studies have shown that it matters whether generation
intervals are measured forward or backward \citep{nishiura2010time,champredon2015intrinsic,britton2019estimation};
we generalize these ideas and show that they apply to any other epidemiological distributions.
Some studies during the early phases of the COVID-19 epidemics have tried to correct for the backward biases, but changes in the backward delay distributions due to changing cohort sizes are expected to be a pervasive feature of outbreak dynamics.
Cohorting epidemiological delays by the primary event time can help avoid backward biases (although censoring biases can still exist) as well as detect potential changes in the distribution.

Here, we assume that all individuals develop symptoms and that the entire transmission process, including all relevant epidemiological delays, is known exactly.
In practice, identifying who infected whom is difficult in general, and asymptomatic and presymptomatic transmission of COVID-19 exacerbates this difficulty \citep{bai2020presumed,he2020temporal,wei2020presymptomatic}.
Biases in the observed serial intervals will necessarily bias the estimates of \RR. 
Furthermore, when one of the individuals is asymptomatic, there is no symptom-based serial interval; 
neglecting the time scale asymptomatic transmission may also bias the estimates of \RR \citep{park2020time}.

>>>>

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|l|l|r|}
\hline
Parameter & Values & Source\\
\hline
Mean intrinsic incubation period & 5.5 days & \cite{lauer2020incubation} \\
SD intrinsic incubation period & 2.4 days & \cite{lauer2020incubation} \\
Mean intrinsic generation interval & 5 days & \cite{ferretti2020quantifying} \\
SD intrinsic generation interval & 2 days & \cite{ferretti2020quantifying} \\
\hline
\end{tabular}
\end{center}
\caption{
  \textbf{Parameter values used for simulations.}
The intrinsic generation-interval distribution is parameterized using a log-normal distribution with log mean $\mu_G=1.54$ and log standard deviation $\sigma_G=0.37$.
The intrinsic incubation period distribution is parameterized using a log-normal distribution with log mean $\mu_I=1.62$ and log standard deviation $\sigma_I=0.42$.
The joint probability distribution is modeled using a bivariate log-normal distribution with correlations (on the log scale) $\rho=\{0, 0.25, 0.5, 0.75\}$.
The intrinsic generation-interval and incubation period distributions are chosen to match characteristic of COVID-19 to illustrate realistic magnitudes of time-varying/perspective effects in the current pandemic.
}
\end{table}

@@

We use parameter estimates for COVID-19 to characterize the degree to which the realized serial-interval distribution can change over the course of an epidemic and to evaluate how different definitions of the serial-interval distribution can affect the Euler-Lotka estimates of \Ro.
We further address how the observed serial intervals, measured through contact tracing, are affected by right censoring during an ongoing epidemic and provide a heuristic method for addressing biases that can arise from using serial-interval data to estimate \Ro.
Finally, we analyze serial-interval data from the COVID-19 epidemic in China under our framework.

@@

\fref{rR} shows Euler-Lotka estimates of \Ro\ based on different definitions of the serial interval. 
When the initial forward serial-interval distribution $f_0(\tau)$ is used, estimates (from \eref{Rforward}) exactly match the (correct) generation-interval-based estimates (\eref{Rgen}) for all values of the correlation $\rho$ between the intrinsic incubation period and the intrinsic generation interval (\fref{rR}A).
When the intrinsic distributions are used, however, estimates based on the serial interval (\eref{Rintrinsic}) underestimate \Ro: as $r$ increases, \Rintrinsic saturates and eventually \emph{decreases} due to the increasing inferred importance of negative serial intervals (\fref{rR}B).
While the initial forward serial intervals during the exponential growth phase can also be negative, their effects are appropriately balanced because faster epidemic growth leads to longer serial intervals (and a corresponding lower proportion of negative intervals).

\begin{figure}[!th]
\includegraphics[width=\textwidth]{rR.pdf}
\caption{
\textbf{Estimates of the reproduction number from the exponential growth rate based on serial- and generation-interval distributions.}
(A). The initial forward serial-interval distributions give a correct
link between the exponential growth rate $r$ and the reproduction
number \Ro,
for any correlation $\rho$ between intrinsic incubation period and
intrinsic generation interval of the underlying bivariate log-normal distribution.
(B) The intrinsic serial-interval distributions give an incorrect link between $r$ and \Ro.
(C) The mean initial forward serial interval during the exponential growth phase increases with $r$.
(D) The squared coefficient of variation of the initial forward serial intervals during the exponential growth phase decreases with $r$.
\bmb{x-axes of B and D don't exactly line up (I realize that's \emph{extremely} fussy)}
}
\label{fig:rR}
\end{figure}

Comparing the shapes of the initial forward serial-interval distribution (\eref{initialSI}) and the intrinsic generation-interval distribution allows us to better understand how different forward distributions lead to identical estimates of \Ro.
In general, distributions with higher means and less variability lead to higher \Ro for a given $r$ \citep{wallinga2007generation, weitz2015modeling, park2019practical}.
When incidence is growing exponentially, forward serial intervals have higher means (\fref{rR}C) and squared coefficients of variation (\fref{rR}D) than the intrinsic generation-interval distribution.
The effects of higher means (which increase \Ro) exactly cancel those of higher variability (which decrease \Ro).
On the other hand, \emph{intrinsic} serial intervals (\eref{intrinsicSI}) have the same mean (equal to the mean initial forward serial at $r=0$ in \fref{rR}C) as the intrinsic generation intervals but are more variable (also see squared coefficient of variation of the initial forward serial-interval distribution at $r=0$ in \fref{rR}D); 
therefore, we underestimate \Ro when we use the intrinsic serial-interval distribution.

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=0.95\textwidth]{forward.pdf}
\caption{
\textbf{Epidemiological dynamics and changes in mean forward and backward delay distributions.}
(A) Daily incidence over time.
(B--D) Changes in the mean forward incubation period, generation interval, and serial interval.
(E--G) Changes in the mean backward incubation period, generation interval, and serial interval.
Black lines represent the results of a deterministic simulation.
Orange lines (A) represent the results of 10 stochastic simulations.
Orange points (B-G) represent the average of 10 stochastic simulations.
Dashed lines represent the mean initial forward delay.
In order to remove possible transient dynamics (e.g., left-censoring of time delays and initial stochasticity due to low number of infections), we set $t=0$ to the first time point when daily incidence is greater than 100.
Intrinsic incubation periods and intrinsic generation intervals are assumed to be independent of each other for simplicity; see Supplementary Materials for simulations with correlated intrinsic incubation periods and intrinsic generation intervals.
See Table 1 for parameter values.
\bmb{identical plot titles B/E etc could be confusing at first glance?}
}
\label{fig:epi}
\end{center}
\end{figure}

@@

We see similar qualitative patterns in all three backward delays (\fref{epi}E--G; \eref{backward}), because they are predominantly driven by the rate of change in incidence, which in turn affects relative cohort sizes.
When incidence is increasing, individuals are more likely to have been infected recently, and therefore we are more likely to observe shorter intervals (\eref{backexp}).
Similarly, when incidence decreases, we are more likely to observe longer intervals.
Neglecting these changes will bias the inference of intrinsic distributions from observed distributions.

@@

While the qualitative changes in the mean forward and backward serial interval are consistent with our earlier simulations (\fref{epi}), the initial mean forward serial interval (\fref{du}B) appears to be larger than what we calculated based on previously estimated incubation period and generation-interval distributions (\fref{rR}C).
This difference may imply that the incubation period and generation interval (Table 1) were underestimated, as neither study explicitly accounted for the fact that the observed intervals were drawn from the backward distributions and were likely to have been censored.

\fref{du}D shows the cohort-averaged estimates of \Ro, which remain roughly constant until day January 17th and suddenly decreases;
this sudden decrease is due to changes in the forward serial intervals consistent with the dynamics seen in our simulations (\fref{obsrR}).
The cohort-averaged estimates of \Ro based on the early forward serial intervals are also consistent with previous estimates of \Ro of the COVID-19 epidemic in China \citep{majumder2020early, doi:10.1098/rsif.2020.0144}:
$\Ro = 2.6$ (95\% CI: 2.2--3.1) and $\Ro = 3.4$ (95\% CI: 2.7 -- 4.3) based on a doubling period of 8 or 6 days, respectively, using serial-interval data from infectors who developed symptoms by January 17th.
These early cohort-averaged estimates of \Ro are unlikely to be affected by the right-censoring as we expect the degree of right-censoring to be low (\fref{du}A).
Therefore, the original \Ro estimate of 1.32 (95\% CI 1.16-–1.48), which neglects the changes in the forward serial-interval distribution, underestimates \Ro by 50\%--60\%.
This example demonstrates the danger of using the observed serial intervals to calculate the reproduction number without organizing serial intervals into cohorts.

@@

Recently, \cite{ali2020serial} also showed that serial intervals of COVID-19 decreased in China. Analogous to our approach, they also grouped serial intervals by the symptom onset date of infectors across 14-day periods and found that the mean forward serial interval decreased from 7.8 days to 2.6 days.
While they attributed the decrease in serial intervals to reduction of the isolation delay, their regression analysis showed that isolation delays explain only 51.5\% of the variation in serial intervals (they could explain up to 72\% of the variance by including other intervention measures).
Our framework provides an explanation for the remaining variation: changes in the backward incubation period during the decreasing phase of an epidemic also shorten serial intervals due to increased amount of presymptomatic transmission (even in the absence of nonpharmaceutical interventions).
Since isolation delays and other intervention measures affect the amount of onward transmission, and therefore the distribution of realized (forward) generation intervals, it cannot explain all variation in forward serial intervals, which depend on (i) backward incubation period of the infector, (ii) the forward generation interval, and (iii) the forward incubation period of the infectee (\fref{diagram}B).

Our results support the use of serial-interval distributions for calculating the initial \RR during the exponential growth phase,  
but they also reveal gaps in current practices in incorporating serial-interval distributions into outbreak analyses.
For example, \cite{thompson2019improved} recently emphasized the importance of using up-to-date serial-interval data for accurate estimation of time-varying reproduction numbers.
However, our results show that if changes the forward serial interval through time are not accounted for, using up-to-date serial-interval data can actually exacerbate the underestimation of initial \RR.
Future studies should explore how neglecting changes in the forward serial-interval distribution can affect the estimates of \RR beyond the exponential growth phase, and potentially re-assess existing estimates of \RR.
We also suggest that modelers should aim to characterize spatiotemporal variation in forward serial-interval distributions and understand their changes over time.
These modeling approaches should be coupled with epidemiological investigation through contact tracing. 
Going forward, an additional advantage of early, intensive contact tracing of emerging diseases is that it provides the best information to characterize the initial forward serial-interval distribution.

Our study also underlines the fact that the serial-interval distribution depends not only on the generation-interval and incubation-period distributions, but also on their intrinsic correlations.
Here, we use a bivariate lognormal distribution to capture their correlations phenomenologically but their true correlations will depend on viral load dynamics;
\bmb{don't understand the connection between these two clauses?}
nonetheless, we expect our conclusions about decreasing serial intervals of COVID-19 to be robust as individuals with longer incubation periods have a longer time window to transmit before symptom onset.
Otherwise \bmb{??}, the impact of increasing backward incubation periods on the forward serial intervals are likely to be disease specific---for example, we show in Supplementary Materials that the initial forward serial-interval distribution can be equivalent to the intrinsic generation-interval distribution, regardless of the growth rate $r$, due to independence between the incubation period and time from symptom onset to transmission (assuming no transmission occurs during the incubation period).
This \bmb{(?antecedent?)} implies that not considering this correlation can potentially bias the estimates of the intrinsic generation-interval distribution from the serial-interval distribution.
% Previous studies that tried to estimate the generation-interval distributions from the observed serial intervals often ignored the dynamical differences between the realized incubation periods of infectors (backward-looking) and those of infectees (forward looking) (e.g., \cite{klinkenberg2011correlation, ganyani2020estimating}).
Better statistical tools for teasing apart the intrinsic generation-interval distributions from the observed serial intervals are needed.

Finally, we lay out a few practical principles for analyzing and interpreting serial-interval data.
First, serial intervals should be cohorted based on the symptom onset date whenever possible.
Previous studies have often regarded serial intervals as an intrinsic quantity, having the same mean as the intrinsic generation interval \citep{svensson2007note,klinkenberg2011correlation,champredon2018equivalence, britton2019estimation}, but the distribution (and the mean) of observed serial intervals differs from this expectation, and changes through time due to epidemic dynamics.
Second, aggregating serial intervals across different cohorts and epidemic periods should be avoided as they \bmb{??} can be subject to different censoring and epidemiological biases:
Even when \emph{all} realized serial intervals can be observed throughout an unmitigated epidemic, we do not obtain the intrinsic serial interval distribution due to susceptible depletion (\fref{obsrR}).
Third, applying serial interval information across epidemics of a given disease should be done with care, because serial intervals are epidemic-specific, rather than disease-specific.
Finally, serial-interval data should be accompanied by a trajectory of the epidemic curve, whenever possible, to provide epidemiological context.
In practice, these recommendations will sometimes be hard to follow, due to limited data about serial intervals, but these issues should be kept in mind when interpreting serial-interval data and understand transmission dynamics at the individual-level.

More broadly, our study underlines the importance of carefully defining measured epidemiological time distributions. 
Previous studies have shown the importance of forward vs. backward
measurement of generation intervals \citep{nishiura2010time,champredon2015intrinsic,britton2019estimation};
we generalize these ideas and show that they apply to any other epidemiological distributions.
Some studies during the early phases of the COVID-19 epidemics have tried to correct for the backward biases \bmb{refs?}, but changes in the backward delay distributions due to changing cohort sizes are expected to be a pervasive feature of outbreak dynamics.
Cohorting epidemiological delays by the primary event time can help avoid backward biases (although censoring biases can still exist) as well as detect potential changes in the distribution.

Here, we assume that all individuals develop symptoms and that the entire transmission process, including all relevant epidemiological delays, is known exactly.
In practice, identifying who infected whom is difficult in general, and asymptomatic and presymptomatic transmission of COVID-19 exacerbates this difficulty \citep{bai2020presumed,he2020temporal,wei2020presymptomatic}.
Biases in the observed serial intervals will necessarily bias the estimates of \RR. 
Furthermore, when one of the individuals in a transmission pair is asymptomatic, there is no symptom-based serial interval; 
neglecting the time scale of asymptomatic transmission may also bias the estimates of \RR \citep{park2020time}.
